---
title: "Final Project: NEON Mosquito Pathogen Borne Status Data"
author: "Alexia Ford"
date: "4/3/2018"
bibliography: references.bib
csl: ecology.csl
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This report is testing how the effect temperature has on mosquitos I would also like to look at the geographical effect on mosquito-borne pathogen statuses, using data collected and publicly presented by NEON. This report will explain NEON's work/purpose, NEON's design, test a hypothesis, and lastly analyze the results. The National Ecological Observatory Network is an observation platform for discovering, understanding, and forecasting the impacts of climate change, land use change, and invasive species on continental-scale ecology with sites distributed within 20 eco-climatic domains across United States, Alaska, Hawaii, and Puerto Rico using remote sensing technologies [@kampe2010neon]. NEON has allowed for easy accessible data to the public by providing the data the observatory has collected to scientists, land managers, policy makers, educators, and members of the general public interested in addressing these major ecological challenges. The data produced by NEON are strategically aimed at addressing ecological questions. Questions that are analyzing complex ecological processes and relationships that govern large scale patterns consistent with the measuring the drivers of change, biogeochemical cycling, and population and community responses across a variety of ecosystems in the ultimate hopes of expanding our understanding of ecological processes and to improve our ability to forecast these ecological changes [@thorpe2016introduction].

The NEON project is responsible for measuring the causes and effects of environmental change across the United States within varying domains that represent different landforms, vegetation, climate, and ecosystem dynamics. The project is constantly collecting data to provide and statistically represent ecological, physical, and biological variability across the US. NEON has four different site types that the observatory collects data from, these sites include: core terrestrial sites, core aquatic sites, relocatable terrestrial sites, and lastly relocatable aquatic sites. The sites are collected  using five measurement subsystems of the Observatory including: the Airborne Observation Platform (AOP), Aquatic Instrument System (AIS), Aquatic Observation System (AOS), Terrestrial Instrument System (TIS), and Terrestrial Observation System (TOS) [@thorpe2016introduction]. These five systems of measurement have been used to collect endless data and NEON has continued to do so by properly standardizing protocols across all sites and maintaining community best practices. 

While NEON collects a wide range of data from atmospheric to even ecohydrology this report will specifically be looking at NEON's collection of mosquito count/pathogen data and IR biological temperature data. The National Ecological Observatory Network (NEON) monitors mosquito populations across its broad geographical range of sites because of their prevalence in food webs, sensitivity to abiotic factors, and relevance for human health. Mosquitoes (sentinel taxa) are helpful organisms to use in study because they are a diverse family of insects with varying life spans and generation times with wide geographic distributions that allow for comparisons across the continent as well as across major vegetation types. Mosquitoes are of great importance to human health due to their epidemiological significance with associated diseases and their ecological influence of crops and healthy livestock. Mosquitoes are also ideal for evaluating the ecological effects of global change as mosquitoes are influenced by climate, vegetation, and host availability. Mosquitoes generally respond quickly to environmental change. Mosquitoes are highly mobile and move to new areas depending on climatic conditions. For, example, as climate warms some mosquitoes are expanding their geographical ranges and efficiency of transmission, resulting in mosquitoes having more interactions with humans [@hoekman2016design]. With this discovery this project will specifically be analyzing the expansion of geographical ranges and efficiency of trasnmission as climates warm, resulting in mosquitos having more interactions with humans.Taking this idea I would like to further expand on how great an impact recent global warming and increase in temperatures has had on the interactions between mosquitos and humans by testing temperature against mosquito-borne pathogen status. I would like to test the hypothesis that as temperature increases mosquito-borne pathogen status also increases. 

In addition to analyzing the effect temperature has on mosquitos I would also like to look at the geographical effect on mosquito-borne pathogen statuses, such as mosquito migration to rivers and the abundance of mosquitos found near bodies of water as human abundance is also typically higher near water. I will do this by analyzing the sites that have high mosquito abundance/pathogen status. I would like to test the hypothesis that the sites near water will have an increase in mosquito-borne pathogen status. 


NEON mosquito sampling takes place regularly during a time period in which mosquitoes are active, this varies by domain. Sampling typically takes place more frequently at core sites than relocatable sites. Core sites are prioritized because of the long‐term nature of core site sampling. In addition, core sites are in wild land areas and provide baseline data to assess how ecological systems are changing through time [@hoekman2016design]. ????????/


Delete this placeholder text, and replace it with your introduction text. You should include background on the NEON project including proper citations (e.g. [@thorpe2016introduction]) and other relevent background information related to your specific hypotheses. Add about 2-3 pages here. Across the whole manuscript, you should cite at least 20 peer reviewed articles.

# Methods
The first data set used in this report is NEON's count mosquito data. The mosquito NEON data comes from a wide range of NEON sites with varying domains. The sampling design for NEON mosquito count data is continuous across all sites to ensure data is consistent and comparable across time and space. The sampling methods include sampling mosquitoes at NEON sites using CDC CO2 light traps. All traps use CO2 as bait because CO2 attracts mosquitoes as it is a component of vertebrate breath that mosquitoes use to locate hosts. Once the mosquito is attracted to the trap a fan pulls the mosquitoes into a net bag for live storage until the trap is collected by a field technician. The design of the CDC CO2 light trap has remained intact since 1962, maintaining it's efficiency and effectiveness in use as the design has allowed for the removal of additional "trash" insects not swaying or adding to collection of data in anyway [@mcnelly1989cdc].

The second set used in this report is NEON's mosquito-borne pathogen status data. NEON scientists collect the mosquitos as they do in the count mosquito set using CDC CO2 light traps sampled in the field.After field collection, mosquito samples are sent to an external facility in which they are sorted to remove bycatch and taxonomically identified to species and sex (if possible). Once the trap is collected by a technician minimal processing occurs at the Domain lab. The mosquitoes are frozen, transferred and stored in dry sample vials at -80°C to be sent for identification and pathogen testing at a facility. Mosquitoes are identified at external facilities in order to minimize identification errors and accurately identify mosquitoes to species. Identification is based on visual examination of external morphology. Identifications for a subset of difficult taxa are verified by DNA barcoding. After identification, target species are destructively tested for pathogens. 
Mosquito-borne pathogen sampling involves the testing of all or a subset of collected mosquitoes for infection
by viral pathogens by one or more external facilities. 

The third set is IR biological temperature. NEON scientists collect IR biological temperature by measuring the paramaters of both ambient air and biological temperatures. Ambient temperature is measured at tower sites using single and triple redundant aspirated temperatures sensors. On the other hand, biological temperature, also known as surface temperature is measured using IR (infared) tempeature sensors located in the soil array and on the tower infastructure. These infared temperature sensors offer a non-contact method determining surface temperatures and resolve the radiation emitted by an object through a thermopile used to measure the difference between the target temperature and the seconsor body; and secondly a thermistor measure the temperature of the sensor body to reference the target temperature. Biological temperature can be used simutaneously with other collected measurements to analyze additional topics such as plant respiration, evapotranspiration rates, and stomatal conductance. These measurements by NEON have alloted for information on how climate and land changes alter ecosystem respiration, energy balance, and fire risk. 

This report will specifically be analyzing how the distribution of elevation varies across all NEON sites and species using: 



Describe where the data came from, including the sets of sites you are comparing, how the data you are using were collected in the field, some information about those sites and why you chose them, and how you are analyzing them (e.g. using R to visualize the results and the statistical tests you will use).

## Site Locations and Descriptions
"STER" is short for NEON's North Sterling site. "STER" is a relocateable terrestrial site in the located in the central plains of Colorado just outside of Sterling in Northeastern Colorado. This plain is characterized by flat terrain and respresents an agronomic site within an urban area. "STER" is at the edge of a experimental field used for the long-term sustainable Dryland Agroecosystems Project (DAP) initiated in 1985 at three NEON eastern Colorado sites to study the effects of cropping intensity on production, water use effeiceincy, and soil chemical and physical properties. The "STER" vegetation consisted of the no-till cropping systems before establishment. By 1910 the site was under conventional tillage since it was taken from native sod. From 1910 to 1985 the conventional tillage ranged from moldboard plowing in the early years to sweep tillage in the later years. The primary crop was winter winter grown in a wheat-fallow rotation with occasional growth of proso millet before 1985. In 1986 grass stands were established that contain a mixture of perennial species including both warm and cool season grasses. The climate of this site constists of hot and low humidity summers and winters typically around freezing point but can possibly drop lower. "STER" faces occasional hailstorms and thunderstorms during the growing seasons, as well as seasonal high winds and tornados in North Sterling. The site provides an understanding of the effects of both climate change and chemical climate to understand and contrast ecological process with other and within cross-domain analyses. A standard selection of data is collected from this site however data from this location provides a baseline for evaluating changes due to invasive species and infectious disease along both a rural and urban gradient in time or space. Overall, Colorados national average population is growing 3 times the national average with the population expected to surpass 5.7 million by 2030. 

"JERC" the Jones Ecological Research Center NEON site is a relocatable terrestrial site in Southeast Georgia. This site is located in the Joseph Jone Ecological Research Center a 11,000- hectare rserve located within the Lower Costal Plains and Flatwoods areas in southern Georgia. The site has been managed with low intensity, dormant-season prescribed fires for the past 75 years for every 3-4 years. "JERC" primary vegetation consist of a mix of Longlear Pine, various oaks, and wire grass, in addition to a numerous amount of other perennial grasses and forbs.  

## Field Sampling Design

Add text here. Add text here. In this section instead of first person (I/we), use 'NEON', or 'NEON scientists', since you'll just be describing what they did.

## Data Analysis and Statistics
Using mosquito data collected and provided by NEON this report will use R and RStudio to visualize results and test how the distribution of elevation varies across NEON sites and differing species?????. To begin, the mosquito count, mosquito-borne pathogen status, and IR biological temperature NEON data was read into RStudio. R and RStudio were used in conjunction with five packages including ggplot2 and dplyr to analyze the data. The ANOVA and chi-square tests were used in RStudio for plots made to test for the significance of the results found. The data, code, and other files needed to reproduce the analyses are available on GitHub. 

# Results

Add a number of code chunks in the Results section. These should read in, subset and plot the data as needed (no need to save any figures to pdf, since they will be put into the rendered document when you click 'knit'), and, for any hypotheses that you want to test, an appropriate statistical test. You should have at least 5-10 figures and/or tables in this section, which you use to test your hypotheses against the dataset. 

```{r load-packages}
# If you add any additional packages here, make sure they are
# also listed in the DESCRIPTION file
library("dplyr")
library("tidyr")
library("ggplot2")
library("readr")
library("knitr")
library("lubridate")
```

```{r read-in-data}
mos_pathogen_results <- read_csv("NEON_pathogens-mosquito/stackedFiles/mos_pathogenresults.csv")
count_mos_taxonomist_ID <- read_csv("NEON_count-mosquitoes/stackedFiles/mos_expertTaxonomistIDProcessed.csv")
count_mos_trapping <- read_csv("NEON_count-mosquitoes/stackedFiles/mos_trapping.csv")
aggregate_temperature <- read_csv("neon_aggregated_temperature_data/neon_hourly_temp.csv")

```

Figure 1. Boxplot of the Elevation of all NEON sites
```{r combine-datasets}
# boxplot of the elevation of all NEON sites
summarized_plots <- count_mos_trapping %>%
  group_by(plotID) %>%
  summarise(mean_lat = mean(decimalLatitude, na.rm = TRUE),
            mean_long = mean(decimalLongitude, na.rm = TRUE),
            mean_elev = mean(elevation, na.rm = TRUE),
            nlcdClass = first(nlcdClass))

count_mos_all <- count_mos_taxonomist_ID %>%
  left_join(summarized_plots)

ggplot(count_mos_all, aes(y = mean_elev,
                          x = siteID)) +
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Elevation (m)") +
  ggtitle("Elevation in meters of all NEON Sites")

#r statistical test
anova(lm(elevation ~ siteID,
         data = mos_pathogen_results))
```
**ANOVA Test***
ANOVA p-value: 2.2e-16 
This p-value is highly significant because the p-value is less than 0.001. This tells me that the NEON sites do in fact have differing elevations. Since the p value is so small (2.2e-16) there is a very low percentage that what we are observing is random in this box plot and for this reason we can reject the null hypothesis.

Figure 2. Barplot of the Count of Test Results of Mosquitos at all NEON sites
```{r barplot-of-count-of-test-results-at-all-sites}
# count test results of mosquitos at all NEON sites 
ggplot(mos_pathogen_results,
       aes(x = siteID,
           fill = testResult)) +
  geom_bar() +
  ggtitle("Count of Test Results of Mosquitos at all NEON sites") +
  xlab("Site ID") +
  ylab("Count of Test Results") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#r chi square test
chisq.test(table(mos_pathogen_results$siteID, mos_pathogen_results$testResult))
```
**Chi-square Test**
Chi-square p-value: 2.2e-16
This p-value is highly significant because the p-value is less than 0.001. This tells me that the there are in fact significant differences in test results among NEON sites. Since the p value is so small (2.2e-16) there is a very low percentage that what we are observing is random in this bar plot and for this reason we can reject the null hypothesis.

Figure 3. Mean Elevation Table for Six of the NEON sites
```{r elevation-table}
mos_pathogen_results %>%
  group_by(siteID) %>%
  summarize(mean_elev = mean(elevation, na.rm = TRUE)) %>%
  arrange(desc(mean_elev)) %>%
  kable(col.names = c("Site ID", "Mean elevation (m)"))
```

Figure 4. Barplot for the Test Results of Mosquito Species at all NEON sites
```{r barplot-for-test-result-all-NEON-sites}
# make a barplot for the test Results of mosquito species at all NEON sites
mos_pathogen_results$scientificName <-
  as.vector(lapply(strsplit(mos_pathogen_results$deprecatedVialID, 
                            split = "\\.", ), `[`, 4),
       mode = "character")

mos_pathogen_results %>%
  filter(!is.na(scientificName)) %>%
  ggplot(aes(fill = scientificName,
             x = testResult)) +
  geom_bar(position = position_dodge()) +
  ggtitle("Test Results of Mosquito Species at all NEON sites") +
  xlab("Test Result") +
  ylab("Count of Mosquitos")

#r chi square test
chisq.test(table(mos_pathogen_results$siteID,
                 mos_pathogen_results$scientificName))
```
**Chi-square Test**
Chi-square p-value: 2.2e-16
This p-value is highly significant because the p-value is less than 0.001. This tells me that the there are in fact different species among the different NEON sites. Since the p value is so small (2.2e-16) there is a very low percentage that what we are observing is random in this bar plot and for this reason we can reject the null hypothesis.

Figure 5. Map of the Elevation of Pathogen Results at all NEON Sites
```{r map-of-elevation-pathogen-results-all-sites}
usa_map <- map_data("usa")

points_with_lat_long <- mos_pathogen_results %>%
  group_by(siteID) %>%
  summarise(mean_lat = mean(decimalLatitude),
            mean_long = mean(decimalLongitude))

count_of_test_results <- mos_pathogen_results %>%
  group_by(siteID, testResult) %>%
  tally()

neg_tests <- count_of_test_results %>%
  filter(testResult == "Negative")

pos_tests <- count_of_test_results %>%
  filter(testResult == "Positive")

ggplot() +
  ggtitle("Elevation of Pathogen Results at all NEON Sites") +
  geom_polygon(data = usa_map,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "light blue") +
  coord_fixed(1.3) +
  geom_point(data =  points_with_lat_long,
             aes(x = mean_long,
                 y = mean_lat,
                 size = neg_tests$n)) + # number of negative
  geom_point(data =  points_with_lat_long,
             aes(x = mean_long,
                 y = mean_lat,
                 size = c(8, 0, 0, 4, 62, 0)), # number of positive tests
             color = "blue")
```

Figure 6. Temperature across all NEON sites for Summer Months
```{r summer-temperature-all-NEON-sites}
aggregate_temperature %>%
  filter(year == 2017) %>%
  filter(month %in% c(6, 7, 8)) %>%
  group_by(siteID) %>%
  summarize(mean_temp = mean(bioTempMaximum, na.rm = TRUE))
```

Figure 7. Temperature across all NEON sites for Winter Months
```{r winter-temperature-all-NEON-sites}
aggregate_temperature %>%
  filter(year == 2017) %>%
  filter(month %in% c(12, 1, 2)) %>%
  group_by(siteID) %>%
  summarize(mean_temp = mean(bioTempMaximum, na.rm = TRUE))
```

In addition to a minimum of 5-10 figures/tables (and associated captions), you should include sufficient text in this section to describe what your findings were. Remember that in the Results section you just describe what you found, but you don't interpret it - that happens in the Discussion. 2-3 pages.

# Discussion

Discuss the results you found in the context of what you know about the site/sites and the predominant types of ecosystems in those locations. Think also about the community dynamics, different types of species interactions, food webs, trophic transfer, energy flow, etc. Should be around 3-4 pages interpreting your results and considering future directions you could take in analyzing these data.

# Sources Cited

